### 2017-11-17 paper/sections/1_introduction.tex
paper: addresses #27 and #34
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Storage systems use software-based caches extensively to improve performance
but the policies that guide what data to evict and when to evict vary with the
application or use case. For example, caching file system metadata on both
clients and servers reduces the number of remote procedure calls and
drastically improves the performance of create-heavy workloads common in
HPC~\cite{ren:sc2014-indexfs, patil:fast2011-giga+}. But these techniques have
policies for when to evict and what data to evict that are specific to the
application's behavior, so a new use case may prove to be poor a match for the
selected caching
policy~\cite{xiao:socc15-shardfs,brandt:msst2003-lh,sevilla:sc15-mantle,
weil:sc2004-dyn-metadata, weil:osdi2006-ceph}. In this paper, we evaluate a
variety of caching policies using a data management language/policy engine and
arrive at a customized policy that works well for our target application.
Trying general policies and iterating to a very customized solution confirms
that our prototype can adapt to different types of workloads, but more
importantly, the collection of policies we developed were similar to policies
we designed for file system metadata load balancing, which suggests that there
are generalities in data management policy design that can be used across
domains.

@@

Our target application is ParSplice~\cite{perez:jctc20150parsplice}, a
molecular dynamics simulation that uses a hierarchy of caches and a single
persistent key-value store to store both observed minima across a molecule's
equation of motion (EOM) and the hundreds or thousands of partial trajectories
calculated each second during a parallel job.  The fine-grained data annotation
capabilities provided by key-value storage is a natural match for scientific
simulations like ParSplice. But these simulations rely on a mesh-based
decomposition of a physical region and result in millions or billions of mesh
cells, where each cell contains materials, pressures, temperatures and other
characteristics that are required to accurately simulate phenomena of interest.
Unfortunately, this simulations of this size saturate the capcity and bandwidth
capabilities of a single node so we need more effective data management
techniques, such as cache management or load balancing across a cluster.

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/cache-management.png}\\
\caption{Using our data management language and policy engine, we designed a
dynamically sized caching policy (dashed line) for ParSplice. The solution uses
knowledge about the application to detect key access patterns and adjust the
cache accordingly. Compared to existing ParSplice configurations (solid lines),
our solution saves the most memory without sacrificing performance.
\label{fig:cache-management}}
\end{figure}

The biggest challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a ParSplice cache that stores molecule
coordinates is shown in Figure~\ref{fig:cache-management}. The solid lines are
the existing cache management policies, exposed as configurations in ParSplice.
The default ParSplice configuration uses an unlimited sized cache, shown by the
``No Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where memory is precious and a common goal
is to keep memory for such data structures below 3\%\footnote{Empirically, we
find this threshold works well for most applications}.  While users can
configure ParSplice to resize the cache when it reaches a certain threshold,
this solution requires tuning and parameter sweeps to find the optimal cache
size that sacrifices negligible performance.  Furthermore, this threshold
changes with different initial configurations and clusters, as shown by our
keyspace analysis in Section~\S\ref{sec:parsplice-keyspace-analysis}.  The
``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
poorly configured cache can save memory but at the expense of performance.  Our
dynamically sized cache, shown by the dashed line in
Figure~\ref{fig:cache-management}, detects key access patterns and
re-sizes the cache accordingly. The design of this policy is driven by a
detailed analysis of the key-value accesses over the course of a long running
simulation across a variety of initial conditions.  Without tuning or parameter
sweeps, our solution saves more memory than a hand-tuned cache without
sacrificing performance.  More importantly, it works for a variety initial
conditions without changing the policy itself.

% What is Mantle
To design more flexible cache management policies, like our dynamically sized
cache solution in Figure~\ref{fig:cache-management}, we use the data management
language and policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.
This framework allows us to dynamically explore the effects of different
software-defined cache management strategies for the changing key-value
workloads generated by ParSplice.  Mantle was originally touted as a
programmable file system metadata load balancer, but we realize now that the
collection of abstractions designed for file systems was a control plane that
improved metadata access. So in this paper we refer to Mantle as a policy
engine that injects policies written in our data management language directly
into a running service, such as a file system or key-value store, and show how
this approach is useful for reasoning about and designing different cache
management strategies in ParSplice.  Developers write policies for ``when" they
want data moved and ``how much" of the data to move, then the framework
executes these policies whenever a decision needs to be made.  These
abstractions help developers unfamiliar with the domain quickly deploy new
policies {\it or} well-understood techniques from other domains.  We show that
Mantle:

@@

  \item decomposes cache management into independent policies that can be
  dynamically changed, making the problem more manageable and facilitating rapid
  development. Changing the policy in use is critical in applications such as
  ParSplice that have alternating stable and chaotic key access patterns
  over the course of a long-running simulation.

  \item can be used to quickly deploy a variety of cache management strategies,
  ranging from basic algorithms and heuristics to statistical models and machine
  learning.

>>>>

Storage systems use software-based caches extensively to improve performance
but the policies that guide what data to evict and when to evict vary with the
use case. For example, caching file system metadata on clients and servers
reduces the number of remote procedure calls and improves the performance of
create-heavy workloads common in HPC~\cite{ren:sc2014-indexfs,
patil:fast2011-giga+}. But these the policies for when to evict and what data
to evict are specific to the application's behavior so a new use case may prove
to be poor a match for the selected caching
policy~\cite{xiao:socc15-shardfs,brandt:msst2003-lh,sevilla:sc15-mantle,
weil:sc2004-dyn-metadata, weil:osdi2006-ceph}. In this paper, we evaluate a
variety of caching policies using a data management language/policy engine and
arrive at a customized policy that works well for our target application.  This
process of trying general policies and quickly iterating to a customized
solution shows that our prototype can adapt to different of workloads.

@@

Our target application is ParSplice~\cite{perez:jctc20150parsplice}, a
molecular dynamics simulation that uses a hierarchy of caches and a single
persistent key-value store to store both observed minima across a molecule's
equation of motion (EOM) and the hundreds or thousands of partial trajectories
calculated each second during a parallel job.  The fine-grained data annotation
capabilities provided by key-value storage is a natural match for scientific
simulations like ParSplice. But these simulations rely on a mesh-based
decomposition of a physical region and result in millions or billions of mesh
cells, where each cell contains materials, pressures, temperatures and other
characteristics that are required to accurately simulate phenomena of interest.
Unfortunately, this simulations of this size saturate the capacity and bandwidth
capabilities of a single node so we need more effective data management
techniques, such as cache management or load balancing across a cluster.

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/cache-management.png}\\
\caption{Using our data management language and policy engine, we designed a
dynamically sized caching policy (thick line) for ParSplice.  Compared to
existing configurations (thin lines with \(\times\)'s), our solution saves the most
memory without sacrificing performance and works for a variety of inputs.
\label{fig:cache-management}}
\end{figure}

The biggest challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a single cache that stores molecule
coordinates is shown in Figure~\ref{fig:cache-management}, where the thin solid
lines marked with \(\times\)'s are the existing configurations in ParSplice.
The default configuration uses an unlimited sized cache, shown by the ``No
Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where a common goal is to keep memory for
such data structures below 3\%\footnote{Anecdotally, we find this threshold
works well for most applications.  For reference, a 1GB cache for a distributed
file system is too large in LANL deployments.}. Since ParSplice deploys a cache
per 300 worker processes, large simulations will not scale. Users can configure
ParSplice to evict data when the cache reaches a threshold but this solution
requires tuning and parameter sweeps; the ``Cache (too small)" curve in
Figure~\ref{fig:cache-management} shows how a poorly configured cache can save
memory but at the expense of performance.  Even worse, this threshold changes
with different initial configurations and clusters so tuning needs to be done
for all system permutations.  Our dynamically sized cache, shown by the think
line in Figure~\ref{fig:cache-management}, detects key access patterns and
re-sizes the cache accordingly.  Without tuning or parameter sweeps, our
solution saves more memory than a hand-tuned cache with a negligible
performance degradation (within the baseline's standard deviation) and works
for a variety of initial conditions.

% What is Mantle
To design more flexible cache management policies, like our solution in
Figure~\ref{fig:cache-management}, we use the data management language and
policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.  Mantle was
originally touted as a programmable file system metadata load balancer, but we
realize now that the collection of abstractions was a control plane that
improved metadata access. So in this paper we refer to Mantle as a policy
engine that injects policies written in our data management language directly
into a running service, such as a file system or key-value store, and show how
this approach is useful for reasoning about and designing different cache
management strategies in ParSplice.  We show that Mantle:

@@

  \item decomposes cache management into independent policies that can be
  dynamically changed, making the problem more manageable. Even developers
  unfamiliar with the domain can quickly deploy policies.

  \item can be used to deploy a variety of cache management strategies ranging
  from basic algorithms and heuristics to statistical models and machine
  learning.
